---
title: "Long Island Sound"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	echo=FALSE,
	dpi=300
	)
```

#Change over time in Long Island

Since the collections of Addison Verrill in the late 1800's and early 1900's, over 100 years have passed.  In 2017 Dr Leanne Elder and Eric Lazo-Wasem lead a collecting project of Long Island Sound Invertebrates.  The specimens they collected are available in the collection databse.  This provides an oppourtunity to compare the two data sets from the same localities 100 years apart.  

The internship posting for the collection project is below and provides some background to the project. 
http://peabody.yale.edu/yale-students/footsteps-addison-verrill-2107-internship



#IZ Long Island sound datadump, 
Verrills work in long island sound, late 1800's to early 1900's and the recent 2017 long island sound collecting trip data  

All data is from:
(http://peabody.yale.edu/collections/invertebrate-zoology).


#Verrill Data

Below we create the set of data belonging to Verrill, I am calling this dataset "V".


```{r}
library( tidyverse )
library( magrittr )
library( knitr )
library( plotly )


V = read_csv( "LIV.csv" )

nrow(V)
names(V)

```


#Ginsburg Data

Next we have the ginsburg data set, collected in 2017, I am calling this dataset "G".

```{r}

G = read_csv( "LIG.csv" )

nrow(G)
names(G)

```


#So...let the analysis and comparisons commence!

#lat and long data for verrill

```{r}

Vllsummary_stats = data.frame(Description=character(), N=numeric())
Vllsummary_stats %<>% rbind(list( "All Records", nrow(V) ), stringsAsFactors=FALSE )
Vllsummary_stats %<>% rbind(list( "Lat and Lon", sum( 
	!is.na(V$`latitude`) & !is.na(V$`longitude`), 
	stringsAsFactors=FALSE
	) ) )

names(Vllsummary_stats) = c("Description", "N")

kable(Vllsummary_stats)

```

```{r}

Gllsummary_stats = data.frame(Description=character(), N=numeric())
Gllsummary_stats %<>% rbind(list( "All Records", nrow(G) ), stringsAsFactors=FALSE )
Gllsummary_stats %<>% rbind(list( "Lat and Lon", sum( 
	!is.na(G$`latitude`) & !is.na(G$`longitude`), 
	stringsAsFactors=FALSE
	) ) )

names(Gllsummary_stats) = c("Description", "N")

kable(Gllsummary_stats)

```

#Species Present in Verrill data

```{r}

Vspecies_list = table(V$"Present Sci name")
Vspecies_frame = data.frame(Taxon=names(Vspecies_list), N=Vspecies_list)
Vspecies_frame
kable( Vspecies_frame )


```

Verrill had 184 species 

#Species present in Ginsburg Data

```{r}

Gspecies_list = table(G$"Present Sci name")
Gspecies_frame = data.frame(Taxon=names(Gspecies_list), N=Gspecies_list)

Gspecies_frame

kable( Gspecies_frame )




```

188 Species

#What taxon are present in both of these data groups? What are missing?

```{r}

intersect(Gspecies_frame$"Taxon", Vspecies_frame$"Taxon" )

```

50 species are in common 

#Why??

Taxonomy:  E.g. We have split the groups more now than they were split 100 years ago 


#How many Phyla are there, how many in common?


```{r}

library(ggplot2)

T <- rbind(V,G)


VPC <- ggplot(data = T) + 
      geom_bar(mapping = aes(x = Phylum, fill = Collector), position = "dodge")

VPC + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


## Collecting bias

collecting techniques and expertise, e.g. wash technique




```{r}



Vphyla_list = table(V$"Phylum")
Vphyla_frame = data.frame(Phylum=names(Vphyla_list), N=Vphyla_list, stringsAsFactors = FALSE)


VPhyla = c(table(V$`Phylum`))
VPhyla = Phyla[order(VPhyla, decreasing = TRUE)]
rownames(Vphyla_frame) = NULL
kable(VPhyla)




VPC <- ggplot(data = V) + 
      geom_bar(mapping = aes(x = Phylum, fill = Class), position = "dodge")

VPC + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


By 

```{r}

library(ggplot2)

Gphyla_list = table(G$"Phylum")
Gphyla_frame = data.frame(Phylum=names(Gphyla_list), N=Gphyla_list, stringsAsFactors = FALSE)

GPhyla = c(table(G$`Phylum`))
GPhyla = GPhyla[order(GPhyla, decreasing = TRUE)]
rownames(Gphyla_frame) = NULL
kable(GPhyla)

           
GPC <- ggplot(data = G) + 
      geom_bar(mapping = aes(x = Phylum, fill = Class), position = "dodge")

GPC + theme(axis.text.x = element_text(angle = 90, hjust = 1))


```



#Genus that cross over
There are only 57 Genera present in both Verrill and Ginsburg data sets

```{r}

intersect(G$"Genus", V$"Genus" )

genus_intersection_frame = data.frame(intersect(G$"Genus", V$"Genus"))

```



#Genus distint to each data set

82 new genera found in 2017 by Elder (G dataset) that were not observed by Verrill

```{r}

#return observations in Ginsburg, but not in Verrill.
setdiff(G$"Genus", V$"Genus" )

G_only_frame = data.frame(setdiff(G$"Genus", V$"Genus" ))


```

```{r}

##return observations in Verrill, but not in Ginsburg
setdiff(V$"Genus", G$"Genus") 

V_only_frame = data.frame(setdiff(V$"Genus", G$"Genus"))
```

Verrill notes 87 Genera (in modern sense - with names updated) not observed in 2017

#Where?!

Where are our specimens from?  

  - size of dot to show density of sampling 

```{r}
V_ll <- V[ (!is.na(V$"Lat") & !is.na(V$"Long") ), ]
G_ll <- G[ (!is.na(G$"Lat") & !is.na(G$"Long") ), ]

# Open libraries
library(tidyverse)
library(maps)
library(mapdata)
library(ggplot2)
library(magick)

# Load map
w <- map_data("world")

# Set longitude and latitude limits
longitude = c(-180,180)
latitude = c(-180, 180)

# Set map limits
map_limits = coord_cartesian(xlim = c(-71.5, -73.5), ylim = c(40.5, 41.5))

#Create land
LImap <- borders("world", colour="gray50", fill="gray50") # create a layer of borders

# Use ggplot2 to plot data from 2 dataframes
require(ggplot2)

w = 
  ggplot() +
  LImap + 
  map_limits +
  geom_point(data=V_ll, aes(x=`Long`, y=`Lat`, col="Verrill", size=1, label1=`CatFullNumber`)) +
  geom_point(data=G_ll, aes(x=`Long`, y=`Lat`, col="Ginsburg", size=1)) 

ggplotly(w)


###note this has two different sets of lat and long for long island, I need to make two graphs by manipulating the data set! then send to eric to show the impact of automatic georeferencing for faukner island 

  

```

```{r}

Glocality_list = table(G$NearestNamedPlace)
Glocality_frame = data.frame(Taxon=names(Glocality_list), N=Glocality_list)

kable( Glocality_frame )

Vlocality_list = table(V$NearestNamedPlace)
Vlocality_frame = data.frame(Taxon=names(Vlocality_list), N=Vlocality_list)

kable( Vlocality_frame)


```

density map???
https://www.littlemissdata.com/blog/maps

```{r}

# Install the relevant libraries - do this one time
library("lubridate")
library("ggplot2")
library("data.table")
library("ggrepel")
library("dplyr")
library("data.table")
library("tidyverse")


q <- ggplot() +
     LImap + 
     map_limits 


 q + stat_density2d(
    aes(x = Long, y = Lat, fill = ..level.., alpha =..level..),
    size = 0.2, bins = 30, data = V,
    geom = "polygon") +
    geom_point(aes(x=V$`Long`, y=V$`Lat`, stroke = 2), col="red", data=n, size = 1.5) +
    geom_label_repel(aes(x=`Long`, y=`Lat`, label = label),
    data=n,
    family = 'N', 
    size = 3, 
    box.padding = 0.2, point.padding = 0.3,
    segment.color = 'grey50')
    
    
 
 geom_point(aes(x = x, y = y, stroke = 2), colour=col4, data = n, size =1.5) + 
  geom_label_repel(
    aes(x, y, label = label),
    data=n,
    family = 'Times', 
    size = 3, 
    box.padding = 0.2, point.padding = 0.3,
    segment.color = 'grey50') 




```




# Notes 


### Notes Talking to Eric:

- Visually audit verrill data names and compare with Worms and record within the Emu database, 
- Ginsburg data includes everything that was actually collected  

1. in database search for strings that are likely absolute, verify and audit the completeless of the data 
2. higher groups within the species, bias in the data, harder to find? who was a better collecter (verrill or elder)? 
3. ^^ is the data skewwed?
4. Verrill did dredging and Ginsburg data didnt, anything over 2m likely not to be collected by hand 

- alot of specimens do not have depth indicators 
  - go to specific locations that Ginsburg survey included 
  - recreate the datasheets and compare only the same locations sampled 

### New data dump 

Using emu, get a better data dump!   - Auditing data to remove bias 

log in: izi2, password: izinterns2

search summary data (merging and excluding etc where necessarry)
- Long Island Sound, Elder, L. E.
- Long Island Sound, Verrill, A. E.  - fm. (for depth)

- Include phyla and other taxonomic data, location etc that will be useful 
- Exclude specimens that were collected by verrill at depth (search and exclude, "fm.")
- Pull data before doing name check, this name check will need to be done in the csv file so not to edit the database info in Emu.


Name check:  Upload file into R, list by scientfic name and frequency to get a list of names that I will need to cross reference and check.
    - Edit the csv file with any name changes that will need to be noted (use WORMS to check against)
    - May need to make a complete new file to include all the synonyms and alternative names??
    - Check with the file that Eric has 
    
We want INTERTIDAL species ! :)

Of interest:

Catch per unit effort: The number of collecting events, unique locality. Specimens per locality, Elder vs Verrill

Phylum and higher taxa
  - and check references 
  - what is going on ?! woah 
  
With better data, we can more accurrately recereate the ecosystem. 



### Maps!

note that I need to make two graphs by manipulating the data set! then send to eric to show the impact of automatic georef for faukner island:  

- make map with incorrect faukner island data
- correct faukner island long and lat data 
- make map with corrected lat and long of faukner island 

Furthermore:  Make a map with density spread rather than discrete data points

ALSO: fix the captions below and titles above for each graph I have made. 



### What else to do?

? individual vs ecosystem level analysis ?






