---
title: "Long Island Sound"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	echo=FALSE,
	dpi=300
	)
```

#Change over time in Long Island

Since the collections of Addison Verrill in the late 1800's and early 1900's, over 100 years have passed.  In 2017 Dr Leanne Elder and Eric Lazo-Wasem lead a collecting project of Long Island Sound Invertebrates.  The specimens they collected are available in the collection databse.  This provides an oppourtunity to compare the two data sets from the same localities 100 years apart.  

The internship posting for the collection project is below and provides some background to the project. 
http://peabody.yale.edu/yale-students/footsteps-addison-verrill-2107-internship



#IZ Long Island sound datadump, 
Verrills work in long island sound, late 1800's to early 1900's and the recent 2017 long island sound collecting trip data  

All data is from:
(http://peabody.yale.edu/collections/invertebrate-zoology).


#Verrill Data

Below we create the set of data belonging to Verrill, I am calling this dataset "V".


```{r}
library( tidyverse )
library( magrittr )
library( knitr )
library( plotly )


V = read_csv( "LIV.csv")



nrow(V)
names(V)

```


#Ginsburg Data

Next we have the ginsburg data set, collected in 2017, I am calling this dataset "G".

```{r}

G = read_csv( "LIG.csv" )

nrow(G)
names(G)

```


#So...let the analysis and comparisons commence!

#lat and long data for verrill

```{r}

Vllsummary_stats = data.frame(Description=character(), N=numeric())
Vllsummary_stats %<>% rbind(list( "All Records", nrow(V) ), stringsAsFactors=FALSE )
Vllsummary_stats %<>% rbind(list( "Lat and Lon", sum( 
	!is.na(V$`latitude`) & !is.na(V$`longitude`), 
	stringsAsFactors=FALSE
	) ) )

names(Vllsummary_stats) = c("Description", "N")

kable(Vllsummary_stats)

```

```{r}

Gllsummary_stats = data.frame(Description=character(), N=numeric())
Gllsummary_stats %<>% rbind(list( "All Records", nrow(G) ), stringsAsFactors=FALSE )
Gllsummary_stats %<>% rbind(list( "Lat and Lon", sum( 
	!is.na(G$`latitude`) & !is.na(G$`longitude`), 
	stringsAsFactors=FALSE
	) ) )

names(Gllsummary_stats) = c("Description", "N")

kable(Gllsummary_stats)

```

#Species Present in Verrill data

Illustrated below there are 184 unique specimens in the verrill data set 
```{r}

Vspecies_list = table(V$"Present Sci name")
Vspecies_frame = data.frame(Taxon=names(Vspecies_list), N=Vspecies_list)
Vspecies_frame


```


#Species present in Ginsburg Data

As illustrated below there are 188 unique species in the Elder/Ginsburg dataset

```{r}

Gspecies_list = table(G$"Present Sci name")
Gspecies_frame = data.frame(Taxon=names(Gspecies_list), N=Gspecies_list)

Gspecies_frame

```


#What taxon are present in both of these data groups? What are missing?

Using an intersection function I see how many species are present in both data sets:

There are 50 species present in both data sets.

```{r}

intersect(Gspecies_frame$"Taxon", Vspecies_frame$"Taxon" )

```

#Visulized in venn diagram

I visualize the number of unique species present in the data sets in the venn diagram below

```{r}

grid.newpage()
draw.pairwise.venn(188, 184, 50, category = c("Number of Species, Verrill ", 
    "Number of Species, Ginsburg/Elder"), lty = rep("blank", 2), fill = c("light blue", "pink"), alpha = rep(0.5, 2), cat.pos = c(0, 0), cat.dist = rep(0.025, 2))



```


#Why??

Why is the intersection so small ? 

Sampling Bias? 
    - The sampling carried out by Verrill and then in 2017 are likely different, i.e. different expertise,          different methods etc 
    
    - Comparative methodologies 

Taxonomic reasons? 
    - Compared to 130 years ago, today when classifying species scientists have more options of species that        a organism belongs to, i.e. more species are known today than 100 years ago, groups have been further         broken up into subgroups etc.


#Looking at the Phylum level

Below is a bar chart illustrating the the number of specimens(by catalogue number, so not individuals and not species) collected by each of collectors. 

To do this I bound the the two datasets together so I could plot all the data and then seperate them by collecter by colour. 

Main take aways:
    - Verrill was a much more prolific collector of annelids
    - The Ginsburg collecting was comparatively heavier on Molluscs and Arthropods than Verrills collecting


This could suggest potenital collection bias/focus/expertise of the two collectors.
    - e.g. wash technique to irritate arthopods from sediment/rock and collect in greater number


```{r}

library(ggplot2)

T <- rbind(V,G)


TP <- ggplot(data = T) + 
      geom_bar(mapping = aes(x = Phylum, fill = Collector), position = "dodge")

TP + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


## Phylum by Class for Ginsburg

```{r}

library(ggplot2)

Gphyla_list = table(G$"Phylum")
Gphyla_frame = data.frame(Phylum=names(Gphyla_list), N=Gphyla_list, stringsAsFactors = FALSE)

GPhyla = c(table(G$`Phylum`))
GPhyla = GPhyla[order(GPhyla, decreasing = TRUE)]
rownames(Gphyla_frame) = NULL
kable(GPhyla)

           
GPC <- ggplot(data = G) + 
      geom_bar(mapping = aes(x = Phylum, fill = Class), position = "dodge")

GPC + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


## Phylum by Class for Verrill

```{r}

library(ggplot2)

Vphyla_list = table(V$"Phylum")
Vphyla_frame = data.frame(Phylum=names(Vphyla_list), N=Vphyla_list, stringsAsFactors = FALSE)

VPhyla = c(table(V$`Phylum`))
VPhyla = VPhyla[order(VPhyla, decreasing = TRUE)]
rownames(Vphyla_frame) = NULL
kable(VPhyla)

           
VPC <- ggplot(data = V) + 
      geom_bar(mapping = aes(x = Phylum, fill = Class), position = "dodge")

VPC + theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

#Genus level comparison

Speceis cross over may be influenced by further refining of groups descriptions, I now take a step back and compare at the Genus level to see if there in more over lap in the data sets. 

    - Find total number of genera in each data set
    - Find intersection 
    - Plot 
    
    - 138 unique genera in Ginsburg dataset 
    - 143 unique genera in Verrill dataset 
    - 56 unigue genera intersect  (note, the data frame has 57 objects but one of them is "NA" likely showing specimens that do not have complete taxonomic classifications?? so am choosing to ignore it this genera level comparison 
    
There are only 56 Genera present in both Verrill and Ginsburg data sets

```{r}

Ggenera_list = table(G$"Genus")
Ggenera_frame = data.frame(Taxon=names(Ggenera_list), N=Ggenera_list)
#Ggenera_frame

Vgenera_list = table(V$"Genus")
Vgenera_frame = data.frame(Taxon=names(Vgenera_list), N=Vgenera_list)
#Vgenera_frame

intersect(G$"Genus", V$"Genus" )
genus_intersection_frame = data.frame(intersect(G$"Genus", V$"Genus"))

grid.newpage()
draw.pairwise.venn(143, 138, 56, category = c("Number of Genera, Verrill ", 
    "Number of Genera, Ginsburg/Elder"), lty = rep("blank", 2), fill = c("light blue", "pink"), alpha = rep(0.5, 2), cat.pos = c(0, 0), cat.dist = rep(0.025, 2))

```


#Genus distint to each data set
82 new genera found in 2017 by Elder (G dataset) that were not observed by Verrill
Verrill notes 87 Genera (in modern sense - with names updated) not observed in 2017

```{r}

#return observations in Ginsburg, but not in Verrill.
#setdiff(G$"Genus", V$"Genus" )

#G_only_frame = data.frame(setdiff(G$"Genus", V$"Genus" ))

```

```{r}

##return observations in Verrill, but not in Ginsburg
#setdiff(V$"Genus", G$"Genus") 

#V_only_frame = data.frame(setdiff(V$"Genus", G$"Genus"))

```

#Does modern classification have anything to do with this? Are there more species today that we re able to distinguish between when identifying specimens?


Making a radial network to illustrate the hierarchical break down across phylum, class, order, family, genus, species for each of the sub-collections.



```{r}

library(data.tree)
library(radialNetworkR)
library(networkD3)






#making a dataframe for the infomation I want


VVtree_frame <- V %>% select('CatFullNumber', 'Phylum', 'Class', 'Order', 'Family', 'Genus', 'Present Sci name')


Vtree_frame <- VVtree_frame[complete.cases(VVtree_frame), ]
str(VVtree_frame)
Vtree_frame <- na.omit(VVtree_frame)

#filter out specimens with imcomplete taxonomy 




Vtree_frame$pathString <- paste("Verrill Specimens", 
                            Vtree_frame$Phylum, 
                            Vtree_frame$Class, 
                            Vtree_frame$Order,
                            Vtree_frame$Family,
                            Vtree_frame$Genus,
                            Vtree_frame$"Present Sci name",
                            sep = "/")

specimens <- as.Node(Vtree_frame)
print(specimens, limit = 1000)



#convert to Node
Vspecimens <- as.Node(Vtree_frame, pathDelimiter = "/")

#plot with networkD3
VspecimensList <- ToListExplicit(Vspecimens, unname = TRUE)
radialNetwork(VspecimensList, fontSize = 5, opacity = 0.9)

diagonalNetwork(VspecimensList, fontSize = 5, opacity = 0.9)





  

```


```{r}


library(data.tree)
library(radialNetworkR)
library(networkD3)






#making a dataframe for the infomation I want


GGtree_frame <- G %>% select('CatFullNumber', 'Phylum', 'Class', 'Order', 'Family', 'Genus', 'Present Sci name')

#filter out specimens with imcomplete taxonomy

Gtree_frame <- GGtree_frame[complete.cases(GGtree_frame), ]
str(GGtree_frame)
Gtree_frame <- na.omit(GGtree_frame)

 




Gtree_frame$pathString <- paste("Ginsburg Specimens", 
                            Gtree_frame$Phylum, 
                            Gtree_frame$Class, 
                            Gtree_frame$Order,
                            Gtree_frame$Family,
                            Gtree_frame$Genus,
                            Gtree_frame$"Present Sci name",
                            sep = "/")

specimens <- as.Node(Gtree_frame)
print(specimens, limit = 1000)



#convert to Node
Gspecimens <- as.Node(Gtree_frame, pathDelimiter = "/")

#plot with networkD3
GspecimensList <- ToListExplicit(Gspecimens, unname = TRUE)
radialNetwork(GspecimensList, fontSize = 5, opacity = 0.9)

diagonalNetwork(GspecimensList, fontSize = 5, opacity = 0.9)




```





#Where?!

Where are our specimens from?  

  - size of dot to show density of sampling 

```{r}
V_ll <- V[ (!is.na(V$"Lat") & !is.na(V$"Long") ), ]
G_ll <- G[ (!is.na(G$"Lat") & !is.na(G$"Long") ), ]

# Open libraries
library(tidyverse)
library(maps)
library(mapdata)
library(ggplot2)
library(magick)

# Load map
w <- map_data("world")

# Set longitude and latitude limits
longitude = c(-180,180)
latitude = c(-180, 180)

# Set map limits
map_limits = coord_cartesian(xlim = c(-71.5, -73.5), ylim = c(40.5, 41.5))

#Create land
LImap <- borders("world", colour="gray50", fill="gray50") # create a layer of borders

# Use ggplot2 to plot data from 2 dataframes
require(ggplot2)

w = 
  ggplot() +
  LImap + 
  map_limits +
  geom_point(data=V_ll, aes(x=`Long`, y=`Lat`, col="Verrill", size=1, label1=`CatFullNumber`)) +
  geom_point(data=G_ll, aes(x=`Long`, y=`Lat`, col="Ginsburg", size=1)) 

ggplotly(w)


###note this has two different sets of lat and long for long island, I need to make two graphs by manipulating the data set! then send to eric to show the impact of automatic georeferencing for faukner island 

  

```


#PLace Names
```{r}

Glocality_list = table(G$NearestNamedPlace)
Glocality_frame = data.frame(Taxon=names(Glocality_list), N=Glocality_list)

kable( Glocality_frame )

Vlocality_list = table(V$NearestNamedPlace)
Vlocality_frame = data.frame(Taxon=names(Vlocality_list), N=Vlocality_list)

kable( Vlocality_frame)


```

density map???
https://www.littlemissdata.com/blog/maps

```{r}

# Install the relevant libraries - do this one time
library("lubridate")
library("ggplot2")
library("data.table")
library("ggrepel")
library("dplyr")
library("data.table")
library("tidyverse")


q <- ggplot() +
     LImap + 
     map_limits 


 q + stat_density2d(
    aes(x = Long, y = Lat, fill = ..level.., alpha =..level..),
    size = 0.2, bins = 30, data = V,
    geom = "polygon") +
    geom_point(aes(x=V$`Long`, y=V$`Lat`, stroke = 2), col="red", data=n, size = 1.5) +
    geom_label_repel(aes(x=`Long`, y=`Lat`, label = label),
    data=n,
    family = 'N', 
    size = 3, 
    box.padding = 0.2, point.padding = 0.3,
    segment.color = 'grey50')
    
    
 
 geom_point(aes(x = x, y = y, stroke = 2), colour=col4, data = n, size =1.5) + 
  geom_label_repel(
    aes(x, y, label = label),
    data=n,
    family = 'Times', 
    size = 3, 
    box.padding = 0.2, point.padding = 0.3,
    segment.color = 'grey50') 




```




# Notes 


### Notes Talking to Eric:

- Visually audit verrill data names and compare with Worms and record within the Emu database, 
- Ginsburg data includes everything that was actually collected  

1. in database search for strings that are likely absolute, verify and audit the completeless of the data 
2. higher groups within the species, bias in the data, harder to find? who was a better collecter (verrill or elder)? 
3. ^^ is the data skewwed?
4. Verrill did dredging and Ginsburg data didnt, anything over 2m likely not to be collected by hand 

- alot of specimens do not have depth indicators 
  - go to specific locations that Ginsburg survey included 
  - recreate the datasheets and compare only the same locations sampled 

### New data dump 

Using emu, get a better data dump!   - Auditing data to remove bias 

log in: izi2, password: izinterns2

search summary data (merging and excluding etc where necessarry)
- Long Island Sound, Elder, L. E.
- Long Island Sound, Verrill, A. E.  - fm. (for depth)

- Include phyla and other taxonomic data, location etc that will be useful 
- Exclude specimens that were collected by verrill at depth (search and exclude, "fm.")
- Pull data before doing name check, this name check will need to be done in the csv file so not to edit the database info in Emu.


Name check:  Upload file into R, list by scientfic name and frequency to get a list of names that I will need to cross reference and check.
    - Edit the csv file with any name changes that will need to be noted (use WORMS to check against)
    - May need to make a complete new file to include all the synonyms and alternative names??
    - Check with the file that Eric has 
    
We want INTERTIDAL species ! :)

Of interest:

Catch per unit effort: The number of collecting events, unique locality. Specimens per locality, Elder vs Verrill

Phylum and higher taxa
  - and check references 
  - what is going on ?! woah 
  
With better data, we can more accurrately recereate the ecosystem. 



### Maps!

note that I need to make two graphs by manipulating the data set! then send to eric to show the impact of automatic georef for faukner island:  

- make map with incorrect faukner island data
- correct faukner island long and lat data 
- make map with corrected lat and long of faukner island 

Furthermore:  Make a map with density spread rather than discrete data points

ALSO: fix the captions below and titles above for each graph I have made. 



### What else to do?

? individual vs ecosystem level analysis ?






